### 2021-10-17 chapter2/chapter.tex
fix more of Nige's comments -- up to page 25. reorder PH-case section and rewrite. add requirements on D, most importanly linearity

Here, the construction of the approximating QBD-RAP is developed intuitively, from a stochastic modelling perspective. The key observation is that, given \(\{\varphi(t)\}\) is constant, then \(\{X(t)\}\) moves deterministically. Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic. We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \cite{ert2006,hstz2016,ert2006,hht2020,mt2021}. As in \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP. Unlike  \cite{bo2013}, however, here we do not take \(\Delta \to 0\). Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small. This necessarily means that the order of the matrix exponential increases.

@@

\paragraph{Modelling the residual time to exit on no change of phase} 
We first consider no change of phase in \([t,t+u]\). Suppose that \(\varphi(t)=i\in\mathcal S_+\) and \(X(t) = x\in\mathcal D_{k,i}\). Observe that on no change of phase, \(\varphi(t+s) = i\), \(s\in[0,u]\), and at time \(t+s\) the level of the fluid queue is given by \(X(t+s) = x+c_i s\), \(s\in[0,u]\). Further, for \(u>(y_{k+1}-x)/c_i\) the level process \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(y_{k+1} - x)/c_i\). Similarly, for \(i\in\mathcal S_-\), \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(x-y_k)/|c_i|\) on the event that there is no change of phase on \([t,t+u]\) for \(u>(x-y_k)/|c_i|\).

To approximate the deterministic behaviour observed above, consider a matrix-exponential distribution \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) with mean \(\Delta\) and low variance. As we shall see later, the assumption that \(Z\) has low variance will allow us to claim that \(Z\) can be used to approximate deterministic behaviour. Let \(Z_i\sim ME(\bs \alpha, \bs S_i, \bs s_i)\) where \(\bs S_i = |c_i|\bs S\), \(\bs s_i=-\bs S_i\bs e\), \(i\in\calS\). The random variable \(Z_i\) has mean \(\Delta/|c_i|\). Further, let \(R_i(u)\) be the residual time, \(R_i(u) = (Z_i-u)1\{Z_i-u>0\}\), \(i\in\calS\). Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process remains in \(i\) for at least this amount of time.

After \(u\in[0,\Delta/|c_i|)\) amount of time has elapsed, on the event that \(\varphi(t+s)=i\), for all \(s\in[0,u)\), then \(X(t+u) = y_k + c_i u\). If the phase remains \(i\) for a further \(\Delta/|c_i|-u\) amount of time, then \(\{X(t)\}\) will leave \(\calD_{k,i}\) at exactly this time. Also at time \(t+u\), given \(Z_i>u\), the distribution of the residual time, \(R_i(u)\), has density 
\[f_{R_i(u)}(r) = \cfrac{\bs\alpha e^{\bs S_i(u+ r) } \bs s_i}{\bs\alpha e^{\bs S_iu } \bs e} = \bs A(t+u) e^{\bs S_i r}  \bs s_i.\]
Here, the event \(Z_i>u\) approximates the event \(X(t+u)\in\calD_{k,i}\) and we want the residual time, \(R_i(u)\), to approximate the time until \(\{X(t)\}\) leaves \(\calD_{k,i}\). That is, we want \(R_i(u)\) to approximate a deterministic random variable at \(\Delta/|c_i| - u\). To this end, we observe that for \(\varepsilon>0\) and \(u<(\Delta-\varepsilon)/|c_i|\), then
\begin{align*}
	&\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|-u) ) 
	\\&= \mathbb P(Z_i\in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
	\\&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2}
\end{align*}
since, by Chebyshev's inequality, \(\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon)) \geq 1- \cfrac{\var(Z)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
when \(\var(Z)\) is small. 
That is, when the variance of \(Z\) (equivalently \(Z_i\)) is low, the residual time \(R_i(u)\) will be concentrated around \(\Delta/|c_i| - u\), as required. Figure \ref{fig:residual distributions} gives an example of a density function for a \(Z_i\) with mean \(\Delta=1\) and \(c_i=1\), as well as the density function of \(R_i(0.3)\) conditional on \(Z_i>0.3\) and \(R_i(0.6)\) conditional on \(Z_i>0.6\), for comparison. Observe that the density of the residual life \(R_i(0.3)\) conditional on \(Z_i>0.3\) is concentrated around \(\Delta-0.3 = 0.7\) and, similarly the density of the residual life \(R_i(0.6)\) conditional in \(Z_i>0.6\) is concentrated around \(\Delta-0.6 = 0.4\).

@@

We can interpret the position of the orbit \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\). This gives a heuristic argument as to how we can model the sojourn times in a given interval \(\mathcal D_{k,i}\) on the event that the phase does not change.

We can apply analogous arguments to heuristically develop a model for the sojourn time of the fluid queue in an interval, \(\calD_{k,i}\), \(i\in\calS_-\), on the event that the phase does not change, given the fluid is \(X(t)=y_{k+1}\). In this case \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).

\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\)} 
%Let \(Z_j\sim ME(\bs\alpha,\bs S_j)\) be a matrix exponential random variable with mean \(\Delta/|c_j|\) and let \(R_j(u)\) be the residual time \(R_j(u) = Z_j-u \mid\{Z_j-u>0\}\). %At time \(t+u\) the value of the orbit is \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\), which can be equivalently written as \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_ju/|c_i| } }{\bs\alpha e^{\bs S_ju/|c_i| } \bs e}\).
Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j-|c_i|u/|c_j|\) occurs, on the event \(Z_j>|c_i|u/|c_j|\). The initial vector associated with the residual time \(R_j(|c_i|u/|c_j|)\) is \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\). This is the initial vector which is obtained if at time \(t\) the phase is \(\varphi(t)=j\) and the orbit is \(\bs A(t)=\bs \alpha,\) so that the QBD-RAP has just entered level \(k\), and there is no change of phase by time \(|c_i|u/|c_j|\). Above, we argued that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) can be seen to correspond to the event that \(X(t+u)\) is \(|c_j|\left(|c_i|u/|c_j|\right) = |c_i|u\) units greater than \(y_k\). 
%We argued earlier that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the position of the orbit on the event that at time \(t\) there is a change of level so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=j\), followed by no change of phase by time \(t+|c_i|u/|c_j|\). The analogous event of the fluid queue is that at time \(t\) the phase is \(\varphi(t)=j\), the fluid level is \(X(t)=y_k\), so that the fluid level has just entered an interval \(\calD_k\) at time \(t\), and there is no change of phase by time \(t+|c_i|u/|c_j|\), in which case the fluid level will be \(X(t+|c_i|u/|c_j|)=y_\ell + |c_j|\left(|c_i|u/|c_j|\right)=y_\ell + |c_i|u\). Therefore, we claim that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the correct position of the orbit to approximate the fact that \(\{X(t)\}\) is \(|c_i|u\) units to the right of the point \(y_k\).

Notice that we can re-write the vector 
\(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) as \(\cfrac{\bs \alpha e^{\bs S_iu } }{\bs \alpha e^{\bs S_iu} \bs e }\). This is the orbit position on the event that at time \(t\) there is a change of level, so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=i\), followed by no change of phase by time \(t+u\). We have argued that this is the orbit position which approximates the fact that \(\{X(t)\}\) is \(|c_i|u\) units greater than \(y_k\).

Therefore, at time \(t+u\), if there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\) then the orbit position should not change as in both cases the orbit position \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) corresponds to an approximation to \(X(t)\) being \(|c_i|u\) units greater than \(y_k\).

@@

Further to this argument, observe that the residual time \(R_j(|c_i|u/|c_j|)\) has density 
\[f_{R_j(|c_i|u/|c_j|)}(r) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}e^{\bs S_j r}\bs s_j,\]
%and corresponds to an orbit position of \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). 
and 
\begin{align}
	&\mathbb P(R_j(|c_i|u/|c_j|) \in ((\Delta-|c_i|u-\varepsilon)/|c_j|, (\Delta-|c_i|u + \varepsilon)/|c_j| )) \nonumber
	\\&=\mathbb P(Z_j\in ((\Delta-\varepsilon)/|c_j|, (\Delta + \varepsilon)/|c_j|)) 
	\\&=\mathbb P(Z\in (\Delta-\varepsilon ,  \Delta + \varepsilon )) 
	% \\&=  \mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon)) 
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
%	\\&\geq \cfrac{1-\cfrac{\var(Z_j)}{\varepsilon^2}}{1} 
	\\&\geq 1-\var(Z)^{1/3} \approx 1.
\end{align}
Hence, when the variance of \(Z\) is low, the residual time, \(R_j(|c_i|u/|c_j|)\), is concentrated around \((\Delta - |c_i|u)/|c_j|\) as required.

Therefore, in our model, upon the event of a change from \(i\in\calS_+\) to \(j\in\calS_+\) at time \(t+u\) the orbit should be \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). That is, the initial condition for the next segment of the orbit process is just the current orbit position and the orbit should start to evolve according to the matrix \(\bs S_j\).

@@

\subsection{Computing \(\bs D\)}
In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020}. For this class of CMEs, we take the index \(p\) to be the order of the representation. Moreover, we take \(p\) to be odd. The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}. Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\). However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.

>>>>

Here, the construction of the approximating QBD-RAP is developed intuitively, from a stochastic modelling perspective. The key observation is that, given \(\{\varphi(t)\}\) is constant, then \(\{X(t)\}\) moves deterministically. Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic. We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \cite{ert2006,hstz2016,ert2006,hht2020,mt2021}. We construct a level process for the QBD-RAP and correspond it to a discretisation the level process of the fluid queue. This is similar to \cite{bo2013} where the level process of their QBD corresponds a discretisation of the fluid level. Unlike \cite{bo2013}, however, here we do not take \(\Delta \to 0\). Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small. % This necessarily means that the order of the matrix exponential increases.

@@

\subsection{Modelling the residual time to exit on no change of phase} 
We first consider no change of phase in \([t,t+u]\). Suppose that \(\varphi(t)=i\in\mathcal S_+\) and \(X(t) = x\in\mathcal D_{k,i}\). Observe that on no change of phase, \(\varphi(t+s) = i\), \(s\in[0,u]\), and at time \(t+s\) the level of the fluid queue is given by \(X(t+s) = x+c_i s\), \(s\in[0,u]\). Further, for \(u>(y_{k+1}-x)/c_i\) the level process \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(y_{k+1} - x)/c_i\). Similarly, for \(i\in\mathcal S_-\), \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(x-y_k)/|c_i|\) on the event that there is no change of phase on \([t,t+u]\) for \(u>(x-y_k)/|c_i|\).

To approximate the deterministic behaviour observed above, consider a matrix-exponential distribution \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) with mean \(\Delta\) and low variance. As we shall see later, the assumption that \(Z\) has low variance will allow us to claim that \(Z\) can be used to approximate deterministic behaviour. Let \(Z_i\sim ME(\bs \alpha, \bs S_i, \bs s_i)\) where \(\bs S_i = |c_i|\bs S\), \(\bs s_i=-\bs S_i\bs e\), \(i\in\calS\). The random variable \(Z_i\) has mean \(\Delta/|c_i|\).

Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process of the QBD-RAP remains in \(i\) for at least this amount of time. A sensible approximation for this deterministic event is to choose the value of the orbit process at time \(t\) to be \(\bs \alpha\), \(\bs A(t)=\bs\alpha\). With this choice, the distribution of time until the next event of the QBD-RAP, on the event that the phase process remains constant will be the distribution of \(Z_i\). Observe, 
\begin{align*}
	\mathbb P(Z_i \in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
	&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2},
\end{align*}
since, by Chebyshev's inequality, \(\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon)) \geq 1- \cfrac{\var(Z)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
when \(\var(Z)\) is small. Hence, when the variance of \(Z\) (equivalently \(Z_i\)) is low, then with high probability the time until the next event of the QBD-RAP, given the phase process remains constant will occur at approximately \(\Delta/|c_i|\).

Now, let \(R_i(u)\) be the residual time, \(R_i(u) = (Z_i-u)1\{Z_i-u>0\}\), \(i\in\calS\). After \(u\in[0,\Delta/|c_i|)\) amount of time has elapsed, on the event that \(\varphi(t+s)=i\), for all \(s\in[0,u)\), then \(X(t+u) = y_k + c_i u\). If the phase remains \(i\) for a further \(\Delta/|c_i|-u\) amount of time, then \(\{X(t)\}\) will leave \(\calD_{k,i}\) at exactly this time. Also at time \(t+u\), given \(Z_i>u\), the distribution of the residual time, \(R_i(u)\), has density 
\[f_{R_i(u)}(r) = \cfrac{\bs\alpha e^{\bs S_i(u+ r) } \bs s_i}{\bs\alpha e^{\bs S_iu } \bs e} = \bs A(t+u) e^{\bs S_i r}  \bs s_i.\]
Here, the event \(Z_i>u\) approximates the event \(X(t+u)\in\calD_{k,i}\) and we want the residual time, \(R_i(u)\), to approximate the time until \(\{X(t)\}\) leaves \(\calD_{k,i}\). That is, we want \(R_i(u)\) to approximate a deterministic random variable at \(\Delta/|c_i| - u\). To this end, we observe that for \(\varepsilon>0\) and \(u<(\Delta-\varepsilon)/|c_i|\), then
\begin{align*}
	&\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|-u) ) 
	\\&= \mathbb P(Z_i\in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
	\\&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2}.
\end{align*}
Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
when \(\var(Z)\) is small. 
That is, when the variance of \(Z\) (equivalently \(Z_i\)) is low, the residual time \(R_i(u)\) will be concentrated around \(\Delta/|c_i| - u\), as required. Figure \ref{fig:residual distributions} gives an example of a density function for a \(Z_i\) with mean \(\Delta=1\) and \(c_i=1\), as well as the density function of \(R_i(0.3)\) conditional on \(Z_i>0.3\) and \(R_i(0.6)\) conditional on \(Z_i>0.6\), for comparison. Observe that the density of the residual life \(R_i(0.3)\) conditional on \(Z_i>0.3\) is concentrated around \(\Delta-0.3 = 0.7\) and, similarly the density of the residual life \(R_i(0.6)\) conditional in \(Z_i>0.6\) is concentrated around \(\Delta-0.6 = 0.4\).

@@

We interpret the position of the orbit \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\). This gives a heuristic argument as to how we can model the sojourn times in a given interval \(\mathcal D_{k,i}\) on the event that the phase does not change.

We can apply analogous arguments to heuristically develop a model for the sojourn time of the fluid queue in an interval, \(\calD_{k,i}\), \(i\in\calS_-\), on the event that the phase does not change, given the fluid is \(X(t)=y_{k+1}\). For \(i\in\calS_-\), the orbit position \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).

\subsection{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\)} 
%Let \(Z_j\sim ME(\bs\alpha,\bs S_j)\) be a matrix exponential random variable with mean \(\Delta/|c_j|\) and let \(R_j(u)\) be the residual time \(R_j(u) = Z_j-u \mid\{Z_j-u>0\}\). %At time \(t+u\) the value of the orbit is \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\), which can be equivalently written as \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_ju/|c_i| } }{\bs\alpha e^{\bs S_ju/|c_i| } \bs e}\).

Let \(E\) be the event that, at time \(t\) the orbit of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = i\in\calS_+\), and there are no events of the QBD-RAP before time \(t+u\). On the event \(E\) the orbit position at time \(t+u\) will be 
\begin{align}
	\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}.\label{eqn: orbit i}
\end{align}
Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=i\) and there are no change of phase by time \(t+u\), then \(X(t+u)=y_{k}+c_iu\). That is, \(X(t+u)\) is \(c_iu\) units from the left-hand edge of \(\calD_k\).

If, at time \(t+u\) there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\), then we need to map the orbit position \(\bs A(t+u^-) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}\) to an orbit position which corresponds to the being \(c_iu\) units from the left-hand edge of \(\calD_k\) in phase \(j\). Call this mapping \(\bs D(i,j)(\cdot)\). As noted in Section \ref{sec: qbd-rap}, the mapping must be linear, 
\(\bs D(i,j)(\bs a) = \bs a \bs D(i,j),\, \bs a \in\mathcal A,\)
for some matrix \(\bs D(i,j)\). The matrix \(\bs D(i,j)\) is a modelling choice.

So that the process is a QBD-RAP, the matrix \(\bs D(i,j)\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D(i,j), \bs S)\) is a representation of a matrix exponential distribution. We would also like the matrix \(\bs D(i,j)\) to have the property \(\bs D(i,j) \bs e = \bs e\) as this will mean that the rate at which a change of phase from \(i\) to \(j\) of the QBD-RAP occurs proportional to 
\[\bs A(t+u) \bs D(i,j)\bs e = 1,\]
which is constant and therefore the distribution of time until a change from phase \(i\) to \(j\) is exponential. We can use this to show, for certain models, the distribution of the phase process of the fluid queue and the distribution of the phase process of the QBD-RAP share the same distribution.

An orbit position which corresponds to the level process of the QBD-RAP beign \(c_iu\) units to the right of \(y_k\) in phase \(j\in\calS_+\) is \(\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e}\). To see this, consider the event that at time \(t\) the orbit process of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = j\in\calS_+\), and there are no events of the QBD-RAP before time \(t+|c_i|u/|c_j|\). On this event, the orbit position at time \(t+u\) will be \(\displaystyle \bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e}\). Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=j\) and there are no change of phase by time \(t+|c_i|u/|c_j|\), then \(X(t+u)=y_{k}+c_j|c_i|u/|c_j| = y_k+c_iu\). That is, \(X(t+u)\) is \(c_iu\) units from the left-hand edge of \(\calD_k\).

Now, observe 
\[\cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e} = \cfrac{\bs \alpha e^{\bs S_iu}}{\bs \alpha e^{\bs S_i u}\bs e},\] 
which is exactly (\ref{eqn: orbit i}), the orbit position on the event \(E\). Hence, it is sensible to choose \(\bs D(i,j)=\bs I\).

% Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j-|c_i|u/|c_j|\) occurs, on the event \(Z_j>|c_i|u/|c_j|\). The initial vector associated with the residual time \(R_j(|c_i|u/|c_j|)\) is \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\). This is the initial vector which is obtained if at time \(t\) the phase is \(\varphi(t)=j\) and the orbit is \(\bs A(t)=\bs \alpha,\) so that the QBD-RAP has just entered level \(k\), and there is no change of phase by time \(|c_i|u/|c_j|\). Above, we argued that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) can be seen to correspond to the event that \(X(t+u)\) is \(|c_j|\left(|c_i|u/|c_j|\right) = |c_i|u\) units greater than \(y_k\). 
% %We argued earlier that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the position of the orbit on the event that at time \(t\) there is a change of level so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=j\), followed by no change of phase by time \(t+|c_i|u/|c_j|\). The analogous event of the fluid queue is that at time \(t\) the phase is \(\varphi(t)=j\), the fluid level is \(X(t)=y_k\), so that the fluid level has just entered an interval \(\calD_k\) at time \(t\), and there is no change of phase by time \(t+|c_i|u/|c_j|\), in which case the fluid level will be \(X(t+|c_i|u/|c_j|)=y_\ell + |c_j|\left(|c_i|u/|c_j|\right)=y_\ell + |c_i|u\). Therefore, we claim that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the correct position of the orbit to approximate the fact that \(\{X(t)\}\) is \(|c_i|u\) units to the right of the point \(y_k\).

% Notice that we can re-write the vector 
% \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) as \(\cfrac{\bs \alpha e^{\bs S_iu } }{\bs \alpha e^{\bs S_iu} \bs e }\). This is the orbit position on the event that at time \(t\) there is a change of level, so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=i\), followed by no change of phase by time \(t+u\). We have argued that this is the orbit position which approximates the fact that \(\{X(t)\}\) is \(|c_i|u\) units greater than \(y_k\).

% Therefore, at time \(t+u\), if there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\) then the orbit position should not change as in both cases the orbit position \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) corresponds to an approximation to \(X(t)\) being \(|c_i|u\) units greater than \(y_k\).

@@

Moreover, the residual time \(R_j(|c_i|u/|c_j|)\) has density 
\[f_{R_j(|c_i|u/|c_j|)}(r) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}e^{\bs S_j r}\bs s_j,\]
%and corresponds to an orbit position of \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). 
and 
\begin{align}
	&\mathbb P(R_j(|c_i|u/|c_j|) \in ((\Delta-|c_i|u-\varepsilon)/|c_j|, (\Delta-|c_i|u + \varepsilon)/|c_j| )) \nonumber
	\\&=\mathbb P(Z_j\in ((\Delta-\varepsilon)/|c_j|, (\Delta + \varepsilon)/|c_j|)) 
	\\&=\mathbb P(Z\in (\Delta-\varepsilon ,  \Delta + \varepsilon )) 
	% \\&=  \mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon)) 
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
%	\\&\geq \cfrac{1-\cfrac{\var(Z_j)}{\varepsilon^2}}{1} 
	\\&\geq 1-\var(Z)^{1/3} \approx 1.
\end{align}
Hence, when the variance of \(Z\) is low, the residual time, \(R_j(|c_i|u/|c_j|)\), is concentrated around \((\Delta - |c_i|u)/|c_j|\), as required.

% Therefore, in our model, upon the event of a change from \(i\in\calS_+\) to \(j\in\calS_+\) at time \(t+u\) the orbit should be \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). That is, the initial condition for the next segment of the orbit process is just the current orbit position and the orbit should start to evolve according to the matrix \(\bs S_j\).

@@

% Let \(\bs \Pi = diag\left(\bs \pi \right)\), where \(\vligne{\pi_k}_{k\in\{1,\dots,p\}}=:\bs \pi=\bs \alpha (-\bs{S})^{-1}/m\) and \(m=\bs \alpha (-\bs{S})^{-1}\bs e\). There is a time-reverse representation of a Phase-type distribution given by \((\widetilde{\bs \alpha}, \widetilde{\bs S}, \widetilde{\bs s})\), where \(\widetilde{\bs \alpha} = m\bs s' \bs \Pi \), \(\widetilde{\bs S} = \bs\Pi^{-1}\bs S'\bs \Pi\) and \(\widetilde{\bs s} = \bs \Pi^{-1}\bs \alpha'/m\) \cite[Page 91]{a2008}.

% Rather than use the residual time \(R_i(u)\) to estimate how far \(X(t)\) is from the edge of an interval, we can use the information contained in \(J(t+u)\). Suppose there is a change of phase of the fluid queue from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\) at which time \(J(t+u)=k\). At a random time before \(t+u\) the Phase-type random variable was initialised. Let \(C_{t+u}\) be the \emph{age} of the Phase-type distribution at time \(t+u\), that is, the time since the Phase-type random variable was initialised. Suppose that \(\varphi(v)=i\), \(v\in [t+u-C_{t+u}, t+u)\), that is, there has been no change of phase of the fluid queue during the current life of the Phase-type random variable. Given \(J(t+u)=k\), we wish to construct an approximation to the distribution of \(C_{t+u}\).
